---
title: "Monthly Forecast"
output:
  html_document: default
  pdf_document: default
---
Load all necessary libraries and data sets.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(ggfortify)
library(vars)
library(forecast)
library(fpp2)
library(lubridate)
library(scales)
library(lemon)
options(scipen=999)
library(GGally)
library(plotly)
library(seasonal)

setwd("~/Ubiqum/Project 4/Task 1/Original Data Set")
power2 <- readRDS("~/Ubiqum/Project 4/Task 1/Original Data Set/power2.rds")
```

I start by creating a data frame containing montly totals in watt-hours for each sub meter as well as a total for the entire household. I then store this as a time-series object.

```{r}
pMonth <- power2 %>%
  group_by(month=floor_date(DateTime, "month")) %>%
  summarize_at(vars(S1_Kitchen:S4_Rest), sum, na.rm = TRUE) %>%
  mutate(total = S1_Kitchen + S2_Laundry + S3_WH_AC + S4_Rest)

pMonthTS <- ts(pMonth, frequency = 12, start = c(2006, 12))
```

We can now plot the time series to see that there is seasonality present in the monthly household consumption totals. We can also make a seasonal plot showing the seasonal variation by year.

```{r}
autoplot(pMonthTS[,"total"]) +
  ggtitle("Total Household Electric Consumption 2006-2010") +
  xlab("Year") +
  ylab("Watt Hours")

ggseasonplot(pMonthTS[,"total"], year.labels=TRUE) +
  ylab("Watt Hours") +
  ggtitle("Seasonal Plot: Total Household Consumption")
```

Next, I create a lag plot as well as an autocorrelation plot to further explore seasonality in the data set.

```{r}
gglagplot(pMonthTS[,6])

ggAcf(pMonthTS[,6]) 
```

Using the tsclean function, we identify and replace any outliers (with suggested values) and missing values so that our model performs better.
```{r}
tsoutliers(pMonthTS[,6])
pMonthClean <- tsclean(pMonthTS[,6])
```

Now, we partition the data into a training and testing set. I want to use a full year as a testing set in order to ensure my forecast for the next 12 months is useful.

Next, I train a linear model on the testing set using trend and seasonality as predictors. I also plot the model against the original data.

```{r}
test1 <- window(pMonthClean,start=c(2006,12), end=c(2009,11))
rest1 <- window(pMonthClean, start=c(2009,11))

LMMonth <- tslm(test1 ~ trend + season)
summary(LMMonth)

autoplot(test1, series="Data") +
  autolayer(fitted(LMMonth), series="Fitted") +
  xlab("Year") + 
  ylab("Watt Hours") +
  ggtitle("Montly Electric Consumption")


```

I also check the residuals and plot the errors of the model.
```{r message=FALSE, warning=FALSE}
accuracy(LMMonth)
checkresiduals(LMMonth)

cbind(Data=pMonthClean, Fitted=fitted(LMMonth)) %>%
  as.data.frame() %>%
  ggplot(aes(x = Data, y = Fitted)) +
  geom_point() +
  ylab("Fitted") + xlab("Actual values") +
  ggtitle("Monthly Consumption - Linear Model Errors") +
  geom_abline(intercept=0, slope=1)
```

Next, I apply the model to forecast testing set and plot the forecast against the acutal values from our testing set. I then examine the performance metrics of the testing and training sets as well as the plot to determine the model's accuracy.

The auto-correlation plot of residuals has a significant peak at lag 8 meaning that there is still some useful information in the data that is not captured by my model.

```{r}
LMTestfcast <- forecast(LMMonth, h=12)

autoplot(LMTestfcast, series="Forecast") +
  autolayer(rest1, series="Observed") +
  ggtitle("Monthly Consumption Forecast on Testing Data") +
  xlab("Year") + ylab("Watt Hours")

```


Finally, I deploy the model on the entire data set to forecast the next 12 months of household energy consumption.
```{r}
LMMonthfinal <- tslm(pMonthClean ~ trend + season)
LMFinalFcast <- forecast(LMMonthfinal, h=12)

autoplot(LMFinalFcast)+
  autolayer(fitted(LMMonthfinal), series="Model") +
  ggtitle("Monthly Consumption Forecast for Next 12 Months") +
  xlab("Year") + ylab("Watt Hours")
```

Now that we have forecasted the next 12 months of energy consumption, and we know that there is trend and seasonality to the data, we can decompose the time-series data to separate out the season and the trend. This plot shows the seasonal component, the trend component, and the remainder component of the time series data set.

We see from the decomposed plot that there is siginficant seasonality, but we also see a decreasing trend which is not as easily visible from the full data set.

```{r}
pMonthD <- decompose(pMonthClean)

pMonthClean %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Decomposed Monthly Electric Consumption in Watt-Hours")
```



