---
title: "SubMeter 3 Time Series Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```
In order to perform a more specific analysis on each submeter, I created separate time series object for each submeter based on a total daily consumption for the full time period of the data set (2006 - 2010).

I then decomposed and plotted each sub-meter time series to see seasonal patterns and trends exist in the data for each submeter. Not surprisingly, each of the sub-meters shows seasonality. Interestingly though, sub-meter 3 is the only submeter that shows an increasing trend.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(vars)
library(forecast)
library(fpp2)
library(lubridate)
library(scales)
library(lemon)
options(scipen=999)
library(GGally)
library(plotly)
library(seasonal)

power2 <- readRDS("~/Ubiqum/Project 4/Task 1/Original Data Set/power2.rds")

pMonth <- power2 %>%
  group_by(month=floor_date(DateTime, "month")) %>%
  summarize_at(vars(S1_Kitchen:S4_Rest), sum, na.rm = TRUE) %>%
  mutate(total = S1_Kitchen + S2_Laundry + S3_WH_AC + S4_Rest)

pDay <- power2 %>%
  group_by(day=floor_date(DateTime, "day")) %>%
  summarize_at(vars(S1_Kitchen:S4_Rest), sum, na.rm = TRUE) %>%
  mutate(total = S1_Kitchen + S2_Laundry + S3_WH_AC + S4_Rest)

pDayTS <- ts(pDay, frequency = 365, start = c(2006,350))
pDayKit <- tsclean(pDayTS[,"S1_Kitchen"])

KitDecomp <- decompose(pDayKit, type="multiplicative")
autoplot(KitDecomp) +
  ggtitle("Decomposed Time Series: Sub-meter 1 - Kitchen")

pDayLau <- tsclean(pDayTS[,"S2_Laundry"])
LaunDecomp <- decompose(pDayLau, type="multiplicative")
autoplot(LaunDecomp) +
  ggtitle("Decomposed Time Series: Sub-meter 2 - Laundry Room")

pDayWH <- tsclean(pDayTS[,"S3_WH_AC"])
WHDecomp <- decompose(pDayWH, type="multiplicative")
autoplot(WHDecomp) +
  ggtitle("Decomposed Time Series: Sub-meter 3 - Water Heater & A/C")

pDayRest <- tsclean(pDayTS[,"S4_Rest"])
RestDecomp <- decompose(pDayRest, type="multiplicative")
autoplot(RestDecomp) +
  ggtitle("Decomposed Time Series: Sub-meter 4 - Rest of House")



#plot just the trend components of each submeter
KitDecomp <- decompose(pDayKit, type="multiplicative")
t1 <- autoplot(KitDecomp$trend) +
  ggtitle("Trend: Sub-meter 1 - Kitchen")

pDayLau <- tsclean(pDayTS[,"S2_Laundry"])
LaunDecomp <- decompose(pDayLau, type="multiplicative")
t2 <- autoplot(LaunDecomp$trend) +
  ggtitle("Trend: Sub-meter 2 - Laundry Room")

pDayWH <- tsclean(pDayTS[,"S3_WH_AC"])
WHDecomp <- decompose(pDayWH, type="multiplicative")
t3 <- autoplot(WHDecomp$trend) +
  ggtitle("Trend: Sub-meter 3 - Water Heater & A/C")

pDayRest <- tsclean(pDayTS[,"S4_Rest"])
RestDecomp <- decompose(pDayRest, type="multiplicative")
t4 <- autoplot(RestDecomp$trend) +
  ggtitle("Trend: Sub-meter 4 - Rest of House")

grid.arrange(t1, t2, t3, t4, ncol = 2)
```

Let's take a closer look at data from sub-meter 3 to better understand what's happening with the energy consumption over time. By comparing the minute-by-minute data from the same week in November of each year. By choosing a time period during winter, we can be realtively certain that the only appliance be used during this time period is the water heater. It would be out of the ordinary for the air conditioner to be used during this time.

From these charts, we can see that over the years, the water heater is activated more frequently and it's using a higher amount of watt-hours each time it's activated. This could mean a few different things. The water heater could be losing efficiency and performing worse over time. However, there could be other explanations. Maybe the family had a houseguest during this time and having an additional person in the house caused the water heater usage to increase. Perhaps the family turned the temperature on their water heater up so it had to work harder to maintain that temperature. 

Becasue we cannot be certain, we should sample other time periods to see if the pattern persists.

```{r}
NovUsage <- power2 %>%
  filter(week==45)

ggplot(NovUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
   labs(title = "November Consumption - Water Heater", 
       y= "Watt-Hours Consumed", x="Day of November")
```
Now, let's look at similar data from another time period. We'll look at January and February data this time. 

In the January and February data, we see that the water heater is at times staying on for much longer periods of time in 2009 and 2010. Also, in 2010, the water heater is using a higher amount of watt-hours of energy (sometimes 30 wh) when it's used. In the earlier years, it's using between 15-20 watt-hours when it's used.

We still cannot rule out a change in the family's habits that causing the increase in the usage of the water heater, however, we can be more certain that there is indeed a pattern of higher consumption over time of the water heater specifically. 

```{r}
JanUsage <- power2 %>%
  filter(week==3)

ggplot(JanUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
  labs(title = "January Consumption - Water Heater", 
       y= "Watt-Hours Consumed", x="Day of January")

FebUsage <- power2 %>%
  filter(week==6)

ggplot(FebUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
   labs(title = "February Consumption - Water Heater", 
       y= "Watt-Hours Consumed", x="Day of February")

```

Let's take a look at some summary data to confirm the information we see in the above visualizations. From this chart of the monthly average watt-hours consumed in 2007-2010, we can see that that average is increasing for sub-meter 3 which includes the water heater and air conditioner. We can also see that the average is decreasing for the other sub-meters.

```{r}
year <- pMonth %>%
  mutate(year = year(month)) %>%
  group_by(year) %>%
  summarize(WH_AC = mean(S3_WH_AC), Laundry = mean(S2_Laundry),
            Kitchen = mean(S1_Kitchen), Rest = mean(S4_Rest)) %>%
  gather(SubMeter, MonthAvg, WH_AC:Rest)

ggplot(subset(year, year>2006), aes(x=year, y=MonthAvg)) +
  geom_line(aes(color=SubMeter), size=2) +
  labs(x="Year", y="Average Watt Hours Consumed",
       title="Monthly Average Watt Hours Consumed by Sub-meter",
       color = "Sub-meter")
```


Now, let's look at the air conditioner. By sampling time periods in the summer, we can try to isolate the usage of the air conditioner and see if there are any interesting consumption patterns. Below are samples from July, August, and September.

The pattern in the summer is harder to see which could be because the energy consumption of the air conditioner is more directly impacted by the weather. In this case, it would be helpful to know what temperature the A/C is set at and what the outside temperature was. One thing looks clear though, the family either didn't have an air conditioner or they didn't use it in 2007.

Based on this analysis though, it seems like the water heater is definitely using more energy and the family should consider alterning their habits (lowering the water heater temperature perhaps) or look into servicing/replacing the water heater to create more efficiency. 

```{r}
JulyUsage <- power2 %>%
  filter(week==28)

ggplot(JulyUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
   labs(title = "July Consumption - Water Heater & A/C", 
       y= "Watt-Hours Consumed", x="Day of July")

AugUsage <- power2 %>%
  filter(week==32)

ggplot(AugUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
   labs(title = "August Consumption - Water Heater & A/C", 
       y= "Watt-Hours Consumed", x="Day of August")

SepUsage <- power2 %>%
  filter(week==38)

ggplot(SepUsage, aes(x=DateTime, y=S3_WH_AC)) +
  geom_line() +
  facet_wrap(~year, scales = "free_x", ncol = 1) +
   labs(title = "September Consumption - Water Heater & A/C", 
       y= "Watt-Hours Consumed", x="Day of September")
```
Now let's look at approximately how much more the WH is costing the homeowner from 2007 compared to 2010. The estimated annual cost is increasing as the consumption increases. There is a 20% increase in the cost from 2007 to 2010. By 2010, Sub-meter 3 is costing approximately $87 more a year than it did in 2007. 

The chart below shows the gradual increase which can be hard to notice month after month, year after year. However, when you compare 2007 to 2010, you can see there is a significant increase in the cost.

```{r}

#create monthly average and estimated annual cost
year <- year %>%
  mutate(kWh = MonthAvg/1000) %>%
  mutate(cost = kWh*.147) %>%
  mutate(annualCost = cost*12)

#plot the estimated annual costs for just S3
year %>% filter(SubMeter == "WH_AC" & year != 2006) %>% 
  ggplot(aes(x = year, y = annualCost)) + 
  geom_line(aes(fill = kWh), color="dark blue", size=1.5) +
  geom_text(aes(label = round(annualCost)), nudge_y = 5, color = "dark blue") +
  labs(x=NULL, y=NULL,
       title= "Estimated Annual Cost of Sub-meter 3",
       subtitle = "Water Heater and Air Conditioner") +
  scale_y_continuous(labels = dollar, breaks = seq(0,600, 25)) +
   theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "gray28"),
        title = element_text(colour = "gray28")) 





```

Now let's look at the total annual estimated energy costs for the entire houseold.

```{r}
annualTot <- year %>%
 group_by(year) %>%
 summarize(total = sum(annualCost)) %>%
  filter(year != 2006)

ggplot(annualTot, aes(x=year, y=total)) +
  geom_line(color="dark blue", size = 2) +
  scale_y_continuous(labels = dollar, breaks = seq(1300, 1500, 25)) +
  geom_text(aes(label = round(total)), nudge_y = 10, color = "dark blue") +
  labs(x=NULL, y=NULL,
       title = "Annual Estimated Household Energy Costs") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "gray"),
        title = element_text(colour = "gray28"))

```

Total minutes used for week 45 for each year.
```{r}

NovUsage1 <- NovUsage %>%
  filter(S3_WH_AC != 0) %>%
  group_by(year) %>%
  summarize(avgWH = median(S3_WH_AC))

NovUsage07 <- NovUsage %>%
  filter(year == 2007 & S3_WH_AC != 0)
#3798 minutes = 63 hours

NovUsage08 <- NovUsage %>%
  filter(year == 2008 & S3_WH_AC != 0)
#3496 minutes = 58 hours

NovUsage09 <- NovUsage %>%
  filter(year == 2009 & S3_WH_AC != 0)
#8034 minutes = 133 hours

NovUsage10 <- NovUsage %>%
  filter(year == 2010 & S3_WH_AC != 0)
#7584 minutes = 126 hours
summary(NovUsage10$S3_WH_AC)
summary(NovUsage07$S3_WH_AC)
```

